---
title: "GGPLOT2 Webinar"
author: "Shane Cone"
date: "3/13/2020"
output: html_document
---

```{r}
library(tidyverse)
dat <- read_csv("C2.1.1CONUSv16.0_2020_p_unit_level_activity_enhanced_v2-tg.csv")
dat_ozone <- read_csv("MLK 2017 AQS data Tidy.csv")

View(dat)
str(dat)
length(dat)

dat_long <- dat %>% 
  select(1:37) %>% 
pivot_longer(7:37, names_to = "Metric", values_to = "Values")

View(dat_long)

dat_long %>% 
  filter(State == "DE", Metric == "Maximum hourly heat input (mmbtu)") %>% 
  group_by(Facility) %>% 
  summarize("Sum of Heat Output" = sum(Values))

dat_long %>% 
  filter(Metric == "Maximum hourly heat input (mmbtu)") %>% 
ggplot(aes(x = State, y = Values)) +
  geom_col()


```


## Grammar of Graphics

```{r} 
#Starting with the AQS data

ggplot(dat_ozone)

ggplot(dat_ozone, aes(x = DateTime, y = o3))
```


This looks like an empy plot, too, but what have we actually accomplished here?

In ggplot2 in R, you build a data visualization piece by piece.The first call just supplies the dataset, but provides no mapping. that is why you get an empty/gray square. However, once we have supplied "aes()", we have mapped our x and y variables. This then also provides the **scale** to ggplot, which is why our x and y axes have both a label, **and** a range! 

However, we don't have any data actually plotted yet! That is because we have not procided a "geom", or a geometric object to map onto the chart. 






